---
title: "ED2 Model Summary and Assessment"
author: "Miriam Johnston"
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: yes
output: 
  pdf_document:
    toc: true
geometry: margin=0.5in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,comment=NA,tidy.opts=list(width.cutoff=60))
```

# ---------------INPUT/OUTPUT SUMMARY---------------
## Directories
```{r, echo=FALSE, comment=NA} 
ED2INdir<-Sys.getenv("ED2INdir")
analydir<-Sys.getenv("analydir")
srcdir<-Sys.getenv("srcdir")

#just for testing:
#ED2INdir="/mnt/odyssey/moorcroftfs5/mjohnston/ED2_timeoutput/Case2/run004/"
#analydir="/mnt/odyssey/moorcroftfs5/mjohnston/ED2_timeoutput/Case2/run004/analy/"
#srcdir="/mnt/odyssey/moorcroftfs5/mjohnston/ED2_timeoutput/Case2/EDc/src/"

source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block1_setwd.R")
kable(df) %>%
    kable_styling(full_width=F,position="left")%>%
    column_spec(1,bold=T,border_right=T,width="10em",background="#C3C3C3") 
```

```{r, echo=FALSE}
source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block2_PrepReadED2IN.R")
```

##Input files
```{r,echo=FALSE,results='markup'}
source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block3_ExtractfromED2IN.R")
  kable(df) %>%
    kable_styling(full_width=F,position="left")%>%
    column_spec(1,bold=T,border_right=T,width="10em")  %>%
    column_spec(2,width="15em") %>%
    column_spec(3,width="20em") %>%
    row_spec(0,bold=T,background="#C3C3C3")
```

##Intended Dates
```{r,echo=FALSE}
#Also uses Block3 R code
kable(df2) %>%
  kable_styling(full_width=F,position="left")%>%
    column_spec(1,width="15em")  %>%
    column_spec(2,width="15em") %>%
    row_spec(0,bold=T,background="#C3C3C3")
```

##Output Files
```{r, echo=FALSE,comment=NA}
source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block5_ExtractFileInfo.R")
kable(filesummary) %>%
  kable_styling(full_width=F,position="left")%>%
    column_spec(1,width="4em")  %>%
    column_spec(2,width="5em") %>%
    row_spec(0,bold=T,background="#C3C3C3")
```

##Plant Functional Types 
###(see ED/src/init/ed_params for PFT parameter values)
```{r, echo=FALSE,comment=NA}
source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block4_ExtractFromedparams.R")
```

# ---------------YEARLY DIAGNOSTICS---------------
## Number of cohorts and patches over time
```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block6_YearlyExtractions.R")
if("Y"%in%tscales){
  plot(timeslist,ncohg,type="o",pch=20,xlab="",ylab="NCOHORTS_GLOBAL")
 }
    ```
    
```{r,echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
#Easier to do only one plot per chunk than to figure out layout
if("Y"%in%tscales){
    plot(timeslist,npatg,type="o",pch=20,xlab="",ylab="NPATCHES_GLOBAL")
    } 
```

##Above-ground biomass
```{r, echo=FALSE,comment=NA, fig.height=4.3, fig.width=6,fig.align='center'}
#AGB
if("Y"%in%tscales){
    for(l in 1:length(pfts_num)){
      if(l==1){
    plot(timeslist,agb[,1],col=colors[l],type="o",pch=20,ylim=c(min(agb),max(agb)),
         ylab="AGB (all pat & coh) [kgC/m2]",xlab="")
      } else {
    lines(timeslist,agb[,l],col=colors[l],type="o",pch=20)
      }
      legend("topleft",lty=1,col=c(colors[1:length(pfts_num)]),legend=pfts_num)
    }
  }
    ```
    
## Leaf Area Index
```{r, echo=FALSE,comment=NA, fig.height=4.3, fig.width=6,fig.align='center'}
#LAI
if("Y"%in%tscales){ 
    for(l in 1:length(pfts_num)){
      if(l==1){
    plot(timeslist,lai[,1],col=colors[l],type="o",pch=20,ylim=c(min(lai),max(lai)),
         ylab="LAI (allpat & coh) [m2/m2]",xlab="")
      } else {
    lines(timeslist,lai[,l],col=colors[l],type="o",pch=20)
      }
      legend("topleft",lty=1,col=c(colors[1:length(pfts_num)]),legend=pfts_num)
    }
      }
```

## Number of plants
```{r, echo=FALSE,comment=NA, fig.height=4.3, fig.width=6,fig.align='center'}
  if("Y"%in%tscales){ 
    for(l in 1:length(pfts_num)){
      if(l==1){
    plot(timeslist,nplant[,1],col=colors[l],type="o",pch=20,ylim=c(min(nplant),max(nplant)),
         ylab="NPLANT (all pat & coh) [plants/m2]",xlab="")
      } else {
    lines(timeslist,nplant[,l],col=colors[l],type="o",pch=20)
      }
      legend("topleft",lty=1,col=c(colors[1:length(pfts_num)]),legend=pfts_num)
    }
      }
```

## Basal area 
```{r,echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
  if("Y"%in%tscales){ 
    
    if(length(yearfilesfull)>5){
      cat("Below: showing basal area size distributions of first year, final year, \n year 1/3 through simulation, and year 2/3 through simulation")
    } else {
      cat("Showing basal area size distributions from all available yearfiles.")
    } 
    
    for(p in 1:length(pfts_num)){
     for(k in 1:length(basal_area)){ #k = years, there should be 4 or 5 of them (first, 1/3, 2/3, last)
       max<-max(unlist(lapply(basal_area,function(x) x[pfts_num[p],,1])))
    if(k==1){
      plot(sizeclasses,basal_area[[k]][pfts_num[p],,1],col=colors[k],type="o",pch=20,
           main=paste("PFT:",pfts_num[p]),xaxt="n",
           ylab="BASAL_AREA [cm2/m2]",xlab="Size Classes (cm)",ylim=c(0,max))
      axis(1,at=1:11,labels=c("0-10","10-20","20-30","30-40","40-50",
                              "50-60","60-70","70-80","80-90","90-100",">100"))
    } else {
      lines(sizeclasses,basal_area[[k]][pfts_num[p],,1],col=colors[k],type="o",pch=20)
    }
       legend("topright",lty=1,col=colors[1:length(basal_area)],legend=subtimeslist)
     }
    }
  }
```
  
# ---------------MONTHLY DIAGNOSTICS---------------

##Number of cohorts and patches over time
```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block7_MonthlyExtractions.R")
if("E"%in%tscales){
  plot(timeslist,ncohg,type="o",pch=20,xlab="",ylab="NCOHORTS_GLOBAL")
 }
    ```
    
```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){
  plot(timeslist,npatg,type="o",pch=20,xlab="",ylab="NPATCHES_GLOBAL")
}
```

##Above-ground biomass
```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){
  for(l in 1:length(pfts_num)){
  if(l==1){
    plot(timeslist,agb[,1],col=colors[l],type="o",pch=20,ylim=c(min(agb),max(agb)),
         ylab="AGB (all pat & coh) [kgC/m2]",xlab="")
  } else {
    lines(timeslist,agb[,l],col=colors[l],type="o",pch=20)
  }
  legend("topleft",lty=1,col=colors[1:length(pfts_num)],legend=pfts_num)
}
}
```

##Leaf area index
```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){
  for(l in 1:length(pfts_num)){
  if(l==1){
    plot(timeslist,lai[,1],col=colors[l],type="o",pch=20,ylim=c(min(lai),max(lai)),
         ylab="LAI (allpat & coh) [m2/m2]",xlab="")
  } else {
    lines(timeslist,lai[,l],col=colors[l],type="o",pch=20)
  }
  legend("topleft",lty=1,col=colors[1:length(pfts_num)],legend=pfts_num)
}
}
```

##Number of plants
```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){
  for(l in 1:length(pfts_num)){
  if(l==1){
    plot(timeslist,nplant[,1],col=colors[l],type="o",pch=20,ylim=c(min(nplant),max(nplant)),
         ylab="NPLANT (all pat & coh) [plants/m2]",xlab="")
  } else {
    lines(timeslist,nplant[,l],col=colors[l],type="o",pch=20)
  }
  legend("topleft",lty=1,col=colors[1:length(pfts_num)],legend=pfts_num)
}
}
```

##Basal area
```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){
  if(length(monthfilesfull)>5){
  cat("Below: showing basal area size distributions of first month, last month, \n month 1/3 through simulation, and month 2/3 through simulation")
  } else {
    cat("Showing basal area size distributions from all available yearfiles.")
  }
  for(p in 1:length(pfts_num)){
  for(k in 1:length(basal_area)){ #k = years, there should be 4 of them (first, 1/3, 2/3, last)
    max<-max(unlist(lapply(basal_area,function(x) x[pfts_num[p],,1])))
    if(k==1){
       plot(sizeclasses,basal_area[[k]][pfts_num[p],,1],col=colors[k],type="o",pch=20,
           main=paste("PFT:",pfts_num[p]),xaxt="n",
           ylab="BASAL_AREA [cm2/m2]",xlab="Size Classes (cm)",ylim=c(0,max))
      axis(1,at=1:11,labels=c("0-10","10-20","20-30","30-40","40-50",
                              "50-60","60-70","70-80","80-90","90-100",">100"))
    } else {
      lines(sizeclasses,basal_area[[k]][pfts_num[p],,1],col=colors[k],type="o",pch=20)
    }
    legend("topright",lty=1,col=colors[1:length(basal_area)],legend=subtimeslist)
  }
}
}
```

##Fluxes of carbon, water, sensible heat - yearly patterns
```{r, echo=FALSE,comment=NA,warning=FALSE,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){
  
  if(length(sub_firstlast)==0){
    cat("No complete years of monthly files available; plotting all months")
  } else if (length(sub_firstlast)==12){
    cat("First full year of monthly files = last full year of monthly files; plotting only 1 year")
  } else {
    cat("Below: showing monthly mean fluxes of the first and last full years of simulation")
  }
  

  if(length(subtimeslistyrs)==24){
    plot(subtimeslistyrs[1:12],mmean_carbon_ac[1:12],type="o",pch=20,xlab="month",ylab="MMEAN_CARBON_AC [umol/m2/s]",
         ylim=c(min(mmean_carbon_ac),max(mmean_carbon_ac)))
    lines(subtimeslistyrs[13:24],mmean_carbon_ac[13:24],type="o",pch=20,col="red")
    legend("topright",col=c("black","red"),legend=c(names(tab[minfullyear]),names(tab[maxfullyear])),lty=1)
  } else {
    plot(subtimeslistyrs,mmean_carbon_ac,type="o",pch=20,xlab="month",ylab="MMEAN_CARBON_AC [umol/m2/s]",
         ylim=c(min(mmean_carbon_ac),max(mmean_carbon_ac)))
       
    legend("topright",col=c("black"),legend=y,lty=1)
  }
    }
 
```

```{r, echo=FALSE,comment=NA,warning=FALSE,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){

  if(length(subtimeslistyrs)==24){
    plot(subtimeslistyrs[1:12],mmean_vapor_ac[1:12],type="o",pch=20,xlab="month",ylab="MMEAN_VAPOR_AC [umol/m2/s]",
         ylim=c(min(mmean_vapor_ac),max(mmean_vapor_ac)))
    lines(subtimeslistyrs[13:24],mmean_vapor_ac[13:24],type="o",pch=20,col="red")
    legend("topright",col=c("black","red"),legend=c(names(tab[minfullyear]),names(tab[maxfullyear])),lty=1)
  } else {
    plot(subtimeslistyrs,mmean_vapor_ac,type="o",pch=20,xlab="month",ylab="MMEAN_VAPOR_AC [umol/m2/s]",
         ylim=c(min(mmean_vapor_ac),max(mmean_vapor_ac)))
       
    legend("topright",col=c("black"),legend=y,lty=1)
  }
    }
 
```

```{r, echo=FALSE,comment=NA,warning=FALSE,fig.height=4.3, fig.width=6,fig.align='center'}
if("E"%in%tscales){

  if(length(subtimeslistyrs)==24){
    plot(subtimeslistyrs[1:12],mmean_sensible_ac[1:12],type="o",pch=20,xlab="month",ylab="MMEAN_SENSIBLE_AC [umol/m2/s]",
         ylim=c(min(mmean_sensible_ac),max(mmean_sensible_ac)))
    lines(subtimeslistyrs[13:24],mmean_sensible_ac[13:24],type="o",pch=20,col="red")
    legend("topright",col=c("black","red"),legend=c(names(tab[minfullyear]),names(tab[maxfullyear])),lty=1)
  } else {
    plot(subtimeslistyrs,mmean_sensible_ac,type="o",pch=20,xlab="month",ylab="MMEAN_SENSIBLE_AC [umol/m2/s]",
         ylim=c(min(mmean_sensible_ac),max(mmean_sensible_ac)))
       
    legend("topright",col=c("black"),legend=y,lty=1)
  }
    }
 
```

# ---------------'INSTANTANEOUS' DIAGNOSTICS---------------

##Fluxes of carbon, water, sensible heat - diurnal patterns
###Legend for the flux plots:
```{r, echo=FALSE,comment=NA, fig.height=4.3, fig.width=6,fig.align='center'}
source("/home/miriam/Dropbox/Harvard/Rpackages/ED2ToolsMRJ/ED2Summary_CompanionFxns/Block8_InstantaneousExtractions.R")
#Make a legend for the flux plots:
if("I"%in%tscales){ 
  if(length(fileselec)!=0){
    for(q in 1:length(filedaysB)){
      if(q==1){
      plot(1:3,rep(q,3),col=colors[q],type="o",lty=lty[q],pch=20,ylim=c(1,length(filedaysB)),
         xlim=c(1,6),xaxt="n",yaxt="n",xlab="",ylab="")
      text(4.2,q,filedaysB[q])
      } else {
        lines(1:3,rep(q,3),col=colors[q],type="o",lty=lty[q],pch=20,ylim=c(1,length(filedaysB)),
         xlim=c(1,6),xaxt="n",yaxt="n",xlab="",ylab="")
      text(4.2,q,filedaysB[q])
      }
    }
  }
  } 
```

```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("I"%in%tscales){
 if(length(fileselec)!=0){
    for(w in 1:ncol(cmat)){
      if(w==1){
    plot(1:dataperday,cmat[,w],type="o",pch=20,col=colors[w],lty=lty[w],
         ylim=c(min(cmat),max(cmat)), xlab="",ylab="AVG_CARBON_AC [umol/m2/s]",
         main=paste(pickyr,", First days of all months",sep=""))
          if(dataperday==24){
            title(xlab="Hour of day (UTC)")
          } else if (dataperday==48){
            title(xlab="Half hour of day (UTC)")
          } else if (dataperday==12){
            title(xlab="2-hour increment of day (UTC)")
          } else {
            title(xlab="Day")
          }
      } else {
        lines(1:dataperday,cmat[,w],type="o",pch=20, col=colors[w],lty=lty[w])
      }
    }
  } 
}
```

```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("I"%in%tscales){
 if(length(fileselec)!=0){
    for(w in 1:ncol(vmat)){
      if(w==1){
    plot(1:dataperday,vmat[,w],type="o",pch=20,col=colors[w],lty=lty[w],
         ylim=c(min(vmat),max(vmat)), xlab="",ylab="AVG_VAPOR_AC [kg/m2/s]",
         main=paste(pickyr,", First days of all months",sep=""))
          if(dataperday==24){
            title(xlab="Hour of day (UTC)")
          } else if (dataperday==48){
            title(xlab="Half hour of day (UTC)")
          } else if (dataperday==12){
            title(xlab="2-hour increment of day (UTC)")
          } else {
            title(xlab="Day")
          }
      } else {
        lines(1:dataperday,vmat[,w],type="o",pch=20, col=colors[w],lty=lty[w])
      }
    }
  } 
}
```

```{r, echo=FALSE,comment=NA,fig.height=4.3, fig.width=6,fig.align='center'}
if("I"%in%tscales){
 if(length(fileselec)!=0){
    for(w in 1:ncol(smat)){
      if(w==1){
    plot(1:dataperday,smat[,w],type="o",pch=20,col=colors[w],lty=lty[w],
         ylim=c(min(smat),max(smat)), xlab="",ylab="AVG_SENSIBLE_AC [W/m2]",
         main=paste(pickyr,", First days of all months",sep=""))
          if(dataperday==24){
            title(xlab="Hour of day (UTC)")
          } else if (dataperday==48){
            title(xlab="Half hour of day (UTC)")
          } else if (dataperday==12){
            title(xlab="2-hour increment of day (UTC)")
          } else {
            title(xlab="Day")
          }
      } else {
        lines(1:dataperday,smat[,w],type="o",pch=20, col=colors[w],lty=lty[w])
      }
    }
  } 
}
```
